Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2017-12-21 LaTeX/Turnover.tex
added information about mixed sex-determination systems in rodents. It seems like the Y causes sex reversed females to have increased fitness

Animals and angiosperms exhibit extremely diverse sex-determination systems \citep[reviewed in][]{Bull:1983vi,Charlesworth:2010it,Beukeboom:2014vb,Bachtrog:2014bx}. 
Among species with genetic sex determination of diploid sexes, some taxa have heterogametic males (XY) and homogametic females (XX), including %non-monotreme? 
mammals and most dioecious plants \citep{Ming:2011iy}; whereas other taxa have homogametic males (ZZ) and heterogametic females (ZW), including Lepidoptera and birds. 
Within several taxa, the chromosome that harbours the master sex-determining region changes. 
For example, transitions of the master sex-determining gene between chromosomes or the evolution of new master sex-determining genes have occurred in Salmonids \citep{Li:2011fm,Yano:2012di}, Diptera \citep{Vicoso:2015hf}, and \textit{Oryzias} \citep{Myosho:2012fv}.
%Presentation at evolution found neo sex chromosome in birds, doesn't seem to be published or bioRxiv yet (consider pers com.), title: A previously unknown neo-sex chromosome mediates plumage divergence and speciation in hybridizing birds Jason Sardell; Elizabeth Cooper; J. Albert C. Uy. I wonder if this is actually a fusion event though?
In addition, many gonochoric clades with genetic sex determination exhibit transitions between male (XY) and female (ZW) heterogamety, including snakes \citep{Gamble2017}, lizards \citep{Ezaz:2009tk}, eight of 26 teleost fish families \citep{Mank:2006bt}, true fruit flies \citep[Tephritids,][]{Vicoso:2015hf}, amphibians \citep{Hillis:1990gu}, the angiosperm genus \textit{Silene} \citep{Slancarova:2013dq}, the angiosperm family \textit{Salicaceae} \citep{Pucholt2015,Pucholt2017} and Coleoptera and Hemiptera \citep[][plate 2]{Beukeboom:2014vb}.
Indeed, in some cases, both male and female heterogametic sex-determination systems can be found in the same species, as exhibited by some cichlid species \citep{Ser:2010iq} and \textit{Rana rugosa} \citep{Ogata:2007jm, Miura2007}. 
In addition, multiple transitions have occurred between genetic and environmental sex-determination systems, e.g., in reptiles and fishes \citep{Conover:1987in,Mank:2006bt,Pokorna:2009ui,Ezaz:2009tk,Pen:2010kk,Holleley:2015hc}.

@@

%IDENTIFICATION OF LOCI IN GENOMIC LOCATIONS THAT IMPLICATE THEM IN SEX-CHROMOSOME TURNOVER
%The hypotheses presented here can be empirically investigated in a similar manner to the idea that transitions between sex-determining systems are favoured by linkage to sexually-antagonistic variation. 
%In the case of sexually-antagonistic variation, one supporting observation is that genes expected to be under sexually-antagonistic selection (e.g., those causing bright male colouration) have been found on recently derived sex chromosomes \citep{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
In support of their role in sex chromosome turnover, genes expected to be under sexually-antagonistic selection (e.g., those causing bright male colouration) have been found on recently derived sex chromosomes \citep{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
Our results show that, if loci experiencing overdominance and/or sexually-antagonistic selection can be identified in close linkage with the ancestral sex-determining locus (rather than only the novel sex-determining locus), then they could also be implicated in driving heterogametic transitions between sex-determination systems. 
%As noted by \citet{vanDoorn:2010hu}, it would be prudent to compare closely related clades in order to determine whether observed polymorphisms predate a transition in sex-determination or arose afterwards. %, particularly because sex-linkage allows sexually-antagonistic selection to maintain polymorphisms under a different and larger parameter space \citep{Rice:1987hs,Jordan:2011fj}. 
In addition, we find that overdominance and/or sexually-antagonistic selection on sex-linked loci can maintain multiple sex-determining alleles (e.g. Figures \ref{fig:freqAll}B and \ref{fig:freqAll}C). 
Within a species, both feminizing and masculinizing alleles have been reported in houseflies \citep{Macdonald1978}, midges \citep{Thompson1971}, frogs \citep{Ogata:2007jm}, cichlid fish \citep{Ser:2010iq}, tilapia \citep{Lee2004}, sea bass \citep{Vandeputte2007}, and lab-strains of Zebrafish \citep{Liew2012,Wilson2014}. 
For example, our results suggest a potential mechanism maintaining multiple sex determining alleles in the platyfish (\textit{Xiphophorus maculatus}), in which X,Y, and W alleles segregate at one locus (or two closely-linked loci) near to potentially sexually-antagonistic genes for pigmentation and sexual maturity \citep{Kallman1965,Kallman1968, Volff2001, Schultheis2006}.

>>>>

Animals and angiosperms exhibit extremely diverse sex-determination systems \citep[reviewed in][]{Bull:1983vi,Charlesworth:2010it,Beukeboom:2014vb,Bachtrog:2014bx}. 
Among species with genetic sex determination of diploid sexes, some taxa have heterogametic males (XY) and homogametic females (XX), including %non-monotreme? 
mammals and most dioecious plants \citep{Ming:2011iy}; whereas other taxa have homogametic males (ZZ) and heterogametic females (ZW), including Lepidoptera and birds. 
Within several taxa, the chromosome that harbours the master sex-determining region changes. 
For example, transitions of the master sex-determining gene between chromosomes or the evolution of new master sex-determining genes have occurred in Salmonids \citep{Li:2011fm,Yano:2012di}, Diptera \citep{Vicoso:2015hf}, and \textit{Oryzias} \citep{Myosho:2012fv}.
%Presentation at evolution found neo sex chromosome in birds, doesn't seem to be published or bioRxiv yet (consider pers com.), title: A previously unknown neo-sex chromosome mediates plumage divergence and speciation in hybridizing birds Jason Sardell; Elizabeth Cooper; J. Albert C. Uy. I wonder if this is actually a fusion event though?
In addition, many gonochoric clades with genetic sex determination exhibit transitions between male (XY) and female (ZW) heterogamety, including snakes \citep{Gamble2017}, lizards \citep{Ezaz:2009tk}, eight of 26 teleost fish families \citep{Mank:2006bt}, true fruit flies \citep[Tephritids,][]{Vicoso:2015hf}, amphibians \citep{Hillis:1990gu}, the angiosperm genus \textit{Silene} \citep{Slancarova:2013dq}, the angiosperm family \textit{Salicaceae} \citep{Pucholt2015,Pucholt2017} and Coleoptera and Hemiptera \citep[][plate 2]{Beukeboom:2014vb}.
Indeed, in some cases, both male and female heterogametic sex-determination systems can be found in the same species, as reported in houseflies \citep{Macdonald1978}, midges \citep{Thompson1971}, frogs \citep{Ogata:2007jm}, cichlid fish \citep{Ser:2010iq}, tilapia \citep{Lee2004}, sea bass \citep{Vandeputte2007}, and lab-strains of Zebrafish \citep{Liew2012,Wilson2014}.
In addition, multiple transitions have occurred between genetic and environmental sex-determination systems, e.g., in reptiles and fishes \citep{Conover:1987in,Mank:2006bt,Pokorna:2009ui,Ezaz:2009tk,Pen:2010kk,Holleley:2015hc}.

@@

%IDENTIFICATION OF LOCI IN GENOMIC LOCATIONS THAT IMPLICATE THEM IN SEX-CHROMOSOME TURNOVER
%The hypotheses presented here can be empirically investigated in a similar manner to the idea that transitions between sex-determining systems are favoured by linkage to sexually-antagonistic variation. 
%In the case of sexually-antagonistic variation, one supporting observation is that genes expected to be under sexually-antagonistic selection (e.g., those causing bright male colouration) have been found on recently derived sex chromosomes \citep{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
In support of their role in sex chromosome turnover, genes expected to be under sexually-antagonistic selection (e.g., those causing bright male colouration) have been found on recently derived sex chromosomes \citep{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
Our results show that, if loci experiencing sexually-antagonistic selection and/or overdominance can be identified in close linkage with the ancestral sex-determining locus (rather than only the novel sex-determining locus), then they could also be implicated in driving heterogametic transitions between sex-determination systems and/or maintaining multiple sex-determining alleles. 
%As noted by \citet{vanDoorn:2010hu}, it would be prudent to compare closely related clades in order to determine whether observed polymorphisms predate a transition in sex-determination or arose afterwards. %, particularly because sex-linkage allows sexually-antagonistic selection to maintain polymorphisms under a different and larger parameter space \citep{Rice:1987hs,Jordan:2011fj}. 
%Within a species, both feminizing and masculinizing alleles have been reported in houseflies \citep{Macdonald1978}, midges \citep{Thompson1971}, frogs \citep{Ogata:2007jm}, cichlid fish \citep{Ser:2010iq}, tilapia \citep{Lee2004}, sea bass \citep{Vandeputte2007}, and lab-strains of Zebrafish \citep{Liew2012,Wilson2014}. 
For example, our results suggest a potential mechanism maintaining multiple sex determining alleles in the platyfish (\textit{Xiphophorus maculatus}), in which X,Y, and W alleles segregate at one locus (or two closely-linked loci) near to potentially sexually-antagonistic genes for pigmentation and sexual maturity \citep{Kallman1965,Kallman1968, Volff2001, Schultheis2006}.
Several rodent species also maintain feminizing alleles along with the ancestral X and Y sex-determination alleles \citep[reviewed in][]{Fredga1994}. 
For example, in nine \textit{Akadon} species, it appears that \textit{sry} expression is suppressed by an autosomal feminizing allele, creating XY females \citep{Bianchi2002,Sanchez2010}, which have increased fitness relative to XX females \citep{Hoekstra2001}. 
In \textit{Mus microtoides}, females can have XX, XX$^\ast$ or X$^\ast$Y genotypes \citep{Veyrunes2010}. 
Previous theory would predict that the X$^\ast$ chromosome (or the autosome it is fused to) harbours female beneficial alleles, driving its spread. 
However, XX and XX$^\ast$ females have similar fitness, whereas X$^\ast$Y female fitness is enhanced \citep{Saunders2014}.
Although Y-linkage of female-beneficial alleles is counterintuitive, our tight linkage model suggests that it can be stably maintained and then favour new feminizing mutations, which is a parsimonious explanation for the spread of feminizing alleles in these rodent species.
